---
title: "Thera Bank"
output: html_notebook
---

Loading all necessary libraries
```{r}
library(psych)
library(readxl)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(randomForest)
library(esquisse)
library(dplyr)
library(ROCR)
library(ineq)
library(InformationValue)
library(corrplot)
library(NbClust)
library(cluster)
library(clValid)
library(factoextra)
```

setting up working directory and loading the data.
```{r}
setwd('D:/Smitayan/PGP BABI')
TheraBank=read_excel('Thera Bank_Personal_Loan_Modelling-dataset-1.xlsx',sheet=2,col_names=T)
```

Reading top 5 rows of the dataset
```{r}
head(TheraBank)
```

Column Names:
```{r}
names(TheraBank)
```

Fixing column names
```{r}
colnames(TheraBank)=make.names(colnames(TheraBank))
```

Data Types
```{r}
str(TheraBank)
```

Summary of data
```{r}
summary(TheraBank)
```

#Data Cleansing: Null treatment, Outlier treatment
```{r}
sapply(TheraBank, function(x) sum(is.na(x)))
```

```{r}
TheraBank$Family.members=replace(TheraBank$Family.members,is.na(TheraBank$Family.members),2)
sapply(TheraBank, function(x) sum(is.na(x)))
```

Converting certain attributes to factors
```{r}
TheraBank$Education=as.factor(TheraBank$Education)
TheraBank$Personal.Loan=as.factor(TheraBank$Personal.Loan)
TheraBank$Securities.Account=as.factor(TheraBank$Securities.Account)
TheraBank$CD.Account=as.factor(TheraBank$CD.Account)
TheraBank$Online=as.factor(TheraBank$Online)
TheraBank$CreditCard=as.factor(TheraBank$CreditCard)
TheraBank$Family.members=as.factor(TheraBank$Family.members)
```


```{r}
str(TheraBank)
```

Checking other statistics of the numerical data
```{r}
describe(TheraBank[-c(1,5,8,10,11,12,13,14)],quant = c(0.02,0.05,0.10,0.25,0.75,0.90,0.95,0.99))
```

Replacing records with negative experience levels with 0 years
```{r}

TheraBank$Experience..in.years.=ifelse(TheraBank$Experience..in.years.<quantile(TheraBank$Experience..in.years.,.02),quantile(TheraBank$Experience..in.years.,.02),TheraBank$Experience..in.years.)
```

EDA: Histograms and Boxplots
```{r}
par(mfrow=c(3,2))
boxplot(TheraBank$Age..in.years.,xlab='Age',horizontal = T,col='orange')
hist(TheraBank$Age..in.years.,xlab = 'Age',main='',col = 'orange')
boxplot(TheraBank$Experience..in.years.,xlab='Experience',horizontal = T,col='orange')
hist(TheraBank$Experience..in.years.,xlab = 'Experience',main=' ',col='orange')
boxplot(TheraBank$Income..in.K.month.,xlab='Income',horizontal = T,col='orange')
hist(TheraBank$Income..in.K.month.,xlab = 'Income',main='',col = 'orange')
boxplot(TheraBank$Mortgage,xlab="Mortgage",horizontal = T,col = 'orange')
hist(TheraBank$Mortgage,xlab = 'Mortgage',main = '',col = 'orange')
boxplot(TheraBank$CCAvg,xlab='Avg Spend',horizontal = T,col = 'orange')
hist(TheraBank$CCAvg,xlab = 'Avg Spend',main = '',col = 'orange')

```

Barplots of factor attributes
```{r}
qplot(TheraBank$Family.members,data = TheraBank,geom = 'bar',xlab = 'Family size',ylab = 'count ')
qplot(TheraBank$Education,data = TheraBank,geom = 'bar',xlab = 'Education',ylab = 'count')
qplot(TheraBank$Personal.Loan,data = TheraBank,geom='bar',xlab = 'Personal Loan',ylab = 'count')
qplot(TheraBank$Securities.Account,data = TheraBank,geom = 'bar',xlab = 'Securities Account',ylab = 'count')
qplot(TheraBank$CD.Account,data = TheraBank,geom = 'bar',xlab = 'CD Account',ylab = 'count')
qplot(TheraBank$Online,data = TheraBank,geom = 'bar',xlab = 'Online',ylab = 'count')
qplot(TheraBank$CreditCard,data = TheraBank,geom = 'bar',xlab = 'Credit Card',ylab = 'count')
```

```{r}
str(TheraBank)
```

Corelation plot
```{r}
corrplot(cor(TheraBank[c(2,3,4,7,9)]))
```

Chisquare tests to determine relation between independent variable and dependent variable
```{r}
chisq.test(TheraBank$Education,TheraBank$Personal.Loan)$p.value
chisq.test(TheraBank$Family.members,TheraBank$Personal.Loan)$p.value
chisq.test(TheraBank$Securities.Account,TheraBank$Personal.Loan)$p.value
chisq.test(TheraBank$CD.Account,TheraBank$Personal.Loan)$p.value
chisq.test(TheraBank$Online,TheraBank$Personal.Loan)$p.value
chisq.test(TheraBank$CreditCard,TheraBank$Personal.Loan)$p.value

```


Some barplots to check the relation between some continuous variables and dependent variable(Personal Income)
```{r}
ggplot(data = TheraBank,aes(x=Education,fill=Personal.Loan))+geom_bar(stat = 'count')+theme_minimal()
```


```{r}
ggplot(data = TheraBank,aes(x=CD.Account,fill=Personal.Loan))+geom_bar(stat = 'count')
```


```{r}
ggplot(data = TheraBank,aes(TheraBank$Age..in.years.,fill=Personal.Loan))+geom_bar()+labs(x='Age in years')
```

```{r}
ggplot(data = TheraBank,aes(TheraBank$Experience..in.years.,fill=Personal.Loan))+geom_bar()+labs(x='Experience in years')
```

```{r}
ggplot(data = TheraBank,aes(TheraBank$Income..in.K.month.,fill=Personal.Loan))+geom_bar()+labs(x='Income')
```


```{r}
ggplot(data = TheraBank,aes(TheraBank$CCAvg,fill=Personal.Loan))+geom_histogram()+labs(x='Average Spend')
```

```{r}
ggplot(data = TheraBank,aes(TheraBank$Mortgage,fill=Personal.Loan))+geom_histogram()+labs(x='Mortgage')
```



Determining number of clusters and scaling the number attributes before clustering
```{r}
TheraBank.num = TheraBank[c(2,3,4,7,9)]
TheraBank.num.scaled = scale(TheraBank.num)
```



NbClust Method:
```{r}
list.methods = c("kl", "ch", "hartigan", "ccc", "scott", "marriot", "trcovw", "tracew", "friedman", "rubin", "cindex", "db", "silhouette", "duda", "pseudot2", "beale", "ratkowsky", "ball", "ptbiserial", "frey", "mcclain", "dunn", "sdindex", "sdbw")
list.distance = c("method","manhattan")

nbctabl = as.data.frame(matrix(ncol = length(list.distance), nrow = length(list.methods)))
names(nbctabl) = list.distance

for(i in 1:length(list.methods)){

nb = NbClust(TheraBank.num.scaled, distance = 'manhattan',
             min.nc = 2, max.nc = 10, method='kmeans',
             index =list.methods[i])
nbctabl[i,2] = nb$Best.nc[1]
nbctabl[i,1] = list.methods[i]
}
```


```{r}
table(nbctabl$manhattan)
```



```{r}

gap_stat = clusGap(TheraBank.num.scaled, FUN = kmeans, K.max = 10, B = 20)

fviz_gap_stat(gap_stat) + theme_minimal() + ggtitle("fviz_gap_stat: Gap Statistic")
```

```{r}
fviz_nbclust(TheraBank.num.scaled, kmeans, method = "silhouette", k.max = 10) + theme_minimal() + ggtitle("The Silhouette Plot")
```

WSS/BSS method
```{r}
wss=c()
bss=c()
for (i in 1:10){
  set.seed(123)
  wss[i]=sum(kmeans(TheraBank.num.scaled,centers = i)$withinss)
  bss[i]=sum(kmeans(TheraBank.num.scaled,centers = i)$betweenss)
}

```

```{r}
ss=data.frame(key=c(1:10),wss,bss)
```

```{r}

ggplot(data = ss,aes(x=key,y=bss))+geom_line(data=ss,stat = 'identity',na.rm=T,col='blue')+geom_point(col='black')+theme_minimal()+labs(x='Number of Clusters(k)',title='BSS Plot')+scale_x_continuous(breaks = ss$key)
```


```{r}
ggplot(data = ss,aes(x=key,y=wss))+geom_line(data=ss,stat = 'identity',na.rm=T,col='blue')+geom_point(col='black')+theme_minimal()+labs(x='Number of Clusters(k)',title='WSS Plot')+scale_x_continuous(breaks = ss$key)
```


#Kmeans Clustering with 3 centres

```{r}
set.seed(1000)
clusters = kmeans(TheraBank.num.scaled,centers = 3,nstart = 1)
```

```{r}
clusplot(TheraBank.num.scaled,clusters$cluster,color = T,labels = T,shade = T,lines = 1)

```


```{r}
TheraBank$cluster=clusters$cluster
```

```{r}
Profiles = aggregate(TheraBank[c(2,3,4,7,9)],by=list(TheraBank$cluster),FUN = mean)
Profiles
```

```{r}
table(TheraBank$cluster,TheraBank$Personal.Loan)
```

```{r}
distmatrix=dist(TheraBank.num.scaled,method = 'manhattan')
cluster2 = hclust(distmatrix,method = 'complete')
plot(cluster2)
rect.hclust(cluster2,k=3,border=2)
```



#Decision Tree
Taking out unwanted variables
```{r}
TheraBank.CART = TheraBank[-c(1,5)]
```

splitting data into Train and test sets
```{r}
set.seed('1234')
nsize = nrow(TheraBank.CART)*.70
trainidx = sample(nrow(TheraBank.CART),size=nsize,replace = F)
traindata = TheraBank.CART[trainidx,]
testdata = TheraBank.CART[-trainidx,]
```

```{r}
r.control= rpart.control(minbucket = 10,cp=0,xval = 10)
DTree = rpart(Personal.Loan~.,data = traindata,control = r.control,method = 'class')
```

```{r}
rpart.plot(DTree)
```

```{r}
print(DTree)
```

```{r}
plotcp(DTree)
```

```{r}
summary(DTree)
```

Model performance measurement using Confusion matrix, AUC, KS and Gini.
```{r}
traindata$predicted.class=predict(DTree,traindata,'class')
traindata$predicted.prob=predict(DTree,traindata,'prob')[,"1"]
```


```{r}
table(traindata$Personal.Loan,traindata$predicted.class)
```

```{r}
TPR=294/(294+35)
TPR
TNR=3162/(3162+9)
TNR
Accuracy=(3162+294)/nrow(traindata)
Accuracy

```


```{r}
testdata$predicted.class=predict(DTree,testdata,'class')
testdata$predicted.prob=predict(DTree,testdata,'prob')[,"1"]
```

```{r}
table(testdata$Personal.Loan,testdata$predicted.class)
```

```{r}
TPR=131/(131+20)
TPR
TNR=1343/(1343+6)
TNR
Accuracy=(1343+131)/nrow(testdata)
Accuracy
```

```{r}
predObj = prediction(testdata$predicted.prob, testdata$Personal.Loan)
perf = performance(predObj, "tpr", "fpr")
plot(perf)

```

```{r}
KS = max(perf@y.values[[1]]-perf@x.values[[1]])
KS
auc = performance(predObj,"auc"); 
auc = as.numeric(auc@y.values)
auc
```

```{r}
gini = ineq(testdata$predicted.prob, type="Gini")
gini
```

RANDOM FOREST

```{r}
set.seed('5678')
nsize = nrow(TheraBank.CART)*.70
trainidx2 = sample(nrow(TheraBank.CART),size=nsize,replace = F)
traindata2 = TheraBank.CART[trainidx2,]
testdata2 = TheraBank.CART[-trainidx2,]
```

Creating initial forest with 501 trees and taking 3 attributes in each split
```{r}
set.seed(1000)
rndfor = randomForest(Personal.Loan~.,data = traindata2,ntree=501,mtry=3,nodesize=10,importance=T)
```

```{r}
plot(rndfor,main='')
```


```{r}
plot(rndfor)
legend("topright", c("OOB", "0", "1"), text.col=1:6, lty=1:3, col=1:3)
```

```{r}
print(rndfor)
```
```{r}
varimp=importance(rndfor)
```

```{r}

```


```{r}
varImpPlot(rndfor, type=1, pch=20, col=1, cex=.5, main='')
```


```{r}
trndfor = tuneRF(x=traindata2[-c(8)],
                 y=traindata2$Personal.Loan,
                 mtryStart = 3,
                 stepFactor = 1.5,
                 ntreeTry = 51,
                 trace = T,
                 plot = T,
                 importance=T,
                 doBest = T)
```

```{r}
finalfor = randomForest(Personal.Loan~.,data = traindata2,ntree=51,mtry=3,nodesize=10,importance=T)
```

```{r}
print(finalfor)
```

```{r}
importance(finalfor)
```


```{r}
traindata2$predicted.class = predict(finalfor,traindata2,'class')
traindata2$predicted.prob = predict(finalfor,traindata2,'prob')[,"1"]
testdata2$predicted.class = predict(finalfor,testdata2,'class')
testdata2$predicted.prob = predict(finalfor,testdata2,'prob')[,"1"]
```

```{r}
table(traindata2$Personal.Loan,traindata2$predicted.class)
```

```{r}
TPR = 325/(325+20)
TPR
TNR = 3154/(3154+1)
TNR
Accuracy = (3154+325)/nrow(traindata2)
Accuracy
```

```{r}
table(testdata2$Personal.Loan,testdata2$predicted.class)
```

```{r}
TPR = 118/(118+17)
TPR
TNR = 1360/(1360+5)
TNR
Accuracy = (1360+118)/nrow(testdata2)
Accuracy
```

```{r}
predObj2 = prediction(testdata2$predicted.prob, testdata2$Personal.Loan)
perf2 = performance(predObj2, "tpr", "fpr")
plot(perf2)
```

```{r}
KS2 = max(perf2@y.values[[1]]-perf2@x.values[[1]])
KS2
auc2 = performance(predObj2,"auc"); 
auc2 = as.numeric(auc2@y.values)
auc2
```

```{r}
gini2 = ineq(testdata2$predicted.prob, type="Gini")
gini2
```

